---
title: "CityDensity"
output: html_document
date: "2023-05-04"
---

## R Markdown
This lines of codes compute for the population density of the cities in the Philipines

#Reading csv
Below code is used to read the csv data
```{r}
pop_data <- read.csv("population.csv")
region_data <- read.csv("regionarea.csv")
```

#Counting the Unique Cities
Below are the lines of codes used to count the number of unique cities per region using unique() function and length() function.
```{r include=FALSE}
cityduplicates_per_region <- aggregate(pop_data$CityProvince, by = list(pop_data$Region), FUN = function(x) length(unique(x)))
names(cityduplicates_per_region) <- c("Region", "NumUniqueCity")
unique_cityduplicates_per_region <- merge(cityduplicates_per_region, region_data)
unique_cityduplicates_per_region
```

#Computation
Below are the lines of code used to:
1) Group the data by region
2) Compute the city population density using the general formula: Population per city/(Area of Region/Number of cities in the Region). This will give a unit of persons per square km
  The numerator in the given code extracts the population of all cities per region. This is divided by the area per region assuming that all cities in that region will have equal land area
```{r include=FALSE}

#NCR
popNCR = pop_data[pop_data$Region=="NATIONAL CAPITAL REGION", c("Province", "CityProvince","Barangay","Population")]
sumNCR = aggregate(Population ~ CityProvince, popNCR, sum)
sumNCR$den = sumNCR$Population/(unique_cityduplicates_per_region[4,3]/unique_cityduplicates_per_region[4,2])
sumNCR

#IX
popIX = pop_data[pop_data$Region=="REGION IX - ZAMBOANGA PENINSULA", c("Province", "CityProvince","Barangay","Population")]
sumIX = aggregate(Population ~ CityProvince, popIX, sum)
sumIX$den = sumIX$Population/(unique_cityduplicates_per_region[10,3]/unique_cityduplicates_per_region[10,2])
sumIX

#V
popV = pop_data[pop_data$Region=="REGION V - BICOL", c("Province", "CityProvince","Barangay","Population")]
sumV = aggregate(Population ~ CityProvince, popV, sum)
sumV$den = sumV$Population/(unique_cityduplicates_per_region[11,3]/unique_cityduplicates_per_region[11,2])
sumV

#VI
popVI = pop_data[pop_data$Region=="REGION VI - WESTERN VISAYAS", c("Province", "CityProvince","Barangay","Population")]
sumVI = aggregate(Population ~ CityProvince, popVI, sum)
sumVI$den = sumVI$Population/(unique_cityduplicates_per_region[12,3]/unique_cityduplicates_per_region[12,2])
sumVI

#VII
popVII = pop_data[pop_data$Region=="REGION VII - CENTRAL VISAYAS", c("Province", "CityProvince","Barangay","Population")]
sumVII = aggregate(Population ~ CityProvince, popVII, sum)
sumVII$den = sumVII$Population/(unique_cityduplicates_per_region[13,3]/unique_cityduplicates_per_region[13,2])
sumVII

#VIII
popVIII = pop_data[pop_data$Region=="REGION VIII - EASTERN VISAYAS", c("Province", "CityProvince","Barangay","Population")]
sumVIII = aggregate(Population ~ CityProvince, popVIII, sum)
sumVIII$den = sumVIII$Population/(unique_cityduplicates_per_region[14,3]/unique_cityduplicates_per_region[14,2])
sumVIII

#X
popX = pop_data[pop_data$Region=="REGION X - NORTHERN MINDANAO", c("Province", "CityProvince","Barangay","Population")]
sumX = aggregate(Population ~ CityProvince, popX, sum)
sumX$den = sumX$Population/(unique_cityduplicates_per_region[15,3]/unique_cityduplicates_per_region[15,2])
sumX

#XI
popXI = pop_data[pop_data$Region=="REGION XI - DAVAO ", c("Province", "CityProvince","Barangay","Population")]
sumXI = aggregate(Population ~ CityProvince, popXI, sum)
sumXI$den = sumXI$Population/(unique_cityduplicates_per_region[16,3]/unique_cityduplicates_per_region[16,2])
sumXI

#XII
popXII = pop_data[pop_data$Region=="REGION XII - SOCCSKSARGEN", c("Province", "CityProvince","Barangay","Population")]
sumXII = aggregate(Population ~ CityProvince, popXII, sum)
sumXII$den = sumXII$Population/(unique_cityduplicates_per_region[17,3]/unique_cityduplicates_per_region[17,2])
sumXII

#XIII
popXIII = pop_data[pop_data$Region=="REGION XIII CARAGA", c("Province", "CityProvince","Barangay","Population")]
sumXIII = aggregate(Population ~ CityProvince, popXIII, sum)
sumXIII$den = sumXIII$Population/(unique_cityduplicates_per_region[18,3]/unique_cityduplicates_per_region[18,2])
sumXIII

popARMM = pop_data[pop_data$Region=="AUTONOMOUS REGION IN MUSLIM MINDANAO (ARMM)", c("Province", "CityProvince","Barangay","Population")]
sumARMM = aggregate(Population ~ CityProvince, popARMM, sum)
sumARMM$den = sumARMM$Population/(unique_cityduplicates_per_region[1,3]/unique_cityduplicates_per_region[1,2])
sumARMM

popCAR = pop_data[pop_data$Region=="CORDILLERA ADMINISTRATIVE REGION", c("Province", "CityProvince","Barangay","Population")]
sumCAR = aggregate(Population ~ CityProvince, popCAR, sum)
sumCAR$den = sumCAR$Population/(unique_cityduplicates_per_region[2,3]/unique_cityduplicates_per_region[2,2])
sumCAR

popMIMAROPA = pop_data[pop_data$Region=="MIMAROPA REGION", c("Province", "CityProvince","Barangay","Population")]
sumMIMAROPA = aggregate(Population ~ CityProvince, popMIMAROPA, sum)
sumMIMAROPA$den = sumMIMAROPA$Population/(unique_cityduplicates_per_region[3,3]/unique_cityduplicates_per_region[3,2])
sumMIMAROPA

popNCR = pop_data[pop_data$Region=="NATIONAL CAPITAL REGION", c("Province", "CityProvince","Barangay","Population")]
sumNCR = aggregate(Population ~ CityProvince, popNCR, sum)
sumNCR$den = sumNCR$Population/(unique_cityduplicates_per_region[4,3]/unique_cityduplicates_per_region[4,2])
sumNCR

popNegros = pop_data[pop_data$Region=="NEGROS ISLAND REGION", c("Province", "CityProvince","Barangay","Population")]
sumNegros = aggregate(Population ~ CityProvince, popNegros, sum)
sumNegros$den = sumNegros$Population/(unique_cityduplicates_per_region[5,3]/unique_cityduplicates_per_region[5,2])
sumNegros

popI = pop_data[pop_data$Region=="REGION I - ILOCOS", c("Province", "CityProvince","Barangay","Population")]
sumI = aggregate(Population ~ CityProvince, popI, sum)
sumI$den = sumI$Population/(unique_cityduplicates_per_region[6,3]/unique_cityduplicates_per_region[6,2])
sumI

popII = pop_data[pop_data$Region=="REGION II - CAGAYAN VALLEY", c("Province", "CityProvince","Barangay","Population")]
sumII = aggregate(Population ~ CityProvince, popII, sum)
sumII$den = sumII$Population/(unique_cityduplicates_per_region[7,3]/unique_cityduplicates_per_region[7,2])
sumII

popIII = pop_data[pop_data$Region=="REGION III - CENTRAL LUZON", c("Province", "CityProvince","Barangay","Population")]
sumIII = aggregate(Population ~ CityProvince, popIII, sum)
sumIII$den = sumIII$Population/(unique_cityduplicates_per_region[8,3]/unique_cityduplicates_per_region[8,2])
sumIII

popIVA = pop_data[pop_data$Region=="REGION IV-A CALABARZON", c("Province", "CityProvince","Barangay","Population")]
sumIVA = aggregate(Population ~ CityProvince, popIVA, sum)
sumIVA$den = sumIVA$Population/(unique_cityduplicates_per_region[9,3]/unique_cityduplicates_per_region[9,2])
sumIVA
```

#Merging
The followng lines of codes merge all the regions in a single table using rbind then sorted using order. The head function was used to identify the top five baranagays with the highest population density
```{r include=FALSE}
merge <- rbind(sumARMM,sumCAR,sumMIMAROPA,sumNCR,sumNegros,sumI,sumII,sumIII,sumIVA, sumV, sumVI, sumVII, sumVIII, sumIX, sumX, sumXI, sumXII, sumXIII)
merge
mergesort = merge[order(merge$den, decreasing=TRUE),]
mergetop5 = head(mergesort, 5)
mergetop5
```

#WriteCsv
The following line of code was used to generate a csv file of the top five baranagays with the highest population density
```{r include=FALSE}
write.csv(mergetop5, file="CityDensity.csv",row.names=FALSE)
```