---
title: "appdensitycalculation"
output: html_document
date: "2023-05-04"
---

## R Markdown
This lines of codes compute for the population density of the barangays in the Philipines

#Reading csv
Below code is used to read the csv data

```{r}
population = read.csv("population.csv")
regionarea = read.csv("regionarea.csv")
```

#Computation
Below are the lines of codes used to:
1) Group the data by region
2) Compute the barangay population density using the general formula: Population per Barangay/(Area of Region/Number of Barangays in the Region). This will give a unit of persons per square km
  The numerator in the given code extracts the population of all barangays per region. This is divided by the area per region assuming that all barangays in that region will have equal land area
  
```{r include=FALSE}
#Region IX
popIX = population[population$Region=="REGION IX - ZAMBOANGA PENINSULA", c("Province","CityProvince","Barangay","Population")]
popIX$den = popIX$Population / (regionarea[10,2]/length(population$Region[population$Region == "REGION IX - ZAMBOANGA PENINSULA"]))
popIX

#Region V
popV = population[population$Region=="REGION V - BICOL", c("Province","CityProvince","Barangay","Population")]
popV$den = popV$Population / (regionarea[11,2]/length(population$Region[population$Region == "REGION V - BICOL"]))
popV

#Region VI
popVI = population[population$Region=="REGION VI - WESTERN VISAYAS", c("Province","CityProvince","Barangay","Population")]
popVI$den = popVI$Population / (regionarea[12,2]/length(population$Region[population$Region == "REGION VI - WESTERN VISAYAS"]))
popVI

#Region VII
popVII = population[population$Region=="REGION VII - CENTRAL VISAYAS", c("Province","CityProvince","Barangay","Population")]
popVII$den = popVII$Population / (regionarea[13,2]/length(population$Region[population$Region == "REGION VII - CENTRAL VISAYAS"]))
popVII

#Region VIII
popVIII = population[population$Region=="REGION VIII - EASTERN VISAYAS", c("Province","CityProvince","Barangay","Population")]
popVIII$den = popVIII$Population / (regionarea[14,2]/length(population$Region[population$Region == "REGION VIII - EASTERN VISAYAS"]))
popVIII

#Region X
popX = population[population$Region=="REGION X - NORTHERN MINDANAO", c("Province","CityProvince","Barangay","Population")]
popX$den = popX$Population / (regionarea[15,2]/length(population$Region[population$Region == "REGION X - NORTHERN MINDANAOL"]))
popX

#Region XI
popXI = population[population$Region=="REGION XI - DAVAO ", c("Province","CityProvince","Barangay","Population")]
popXI$den = popXI$Population / (regionarea[16,2]/length(population$Region[population$Region == "REGION XI - DAVAO "]))
popXI

#Region XII
popXII = population[population$Region=="REGION XII - SOCCSKSARGEN", c("Province","CityProvince","Barangay","Population")]
popXII$den = popXII$Population / (regionarea[17,2]/length(population$Region[population$Region == "REGION XII - SOCCSKSARGEN"]))
popXII

#Region XIII
popXIII = population[population$Region=="REGION XIII CARAGA", c("Province","CityProvince","Barangay","Population")]
popXIII$den = popXIII$Population / (regionarea[18,2]/length(population$Region[population$Region == "REGION XIII CARAGA"]))
popXIII

#Region IV-A
popIVA = population[population$Region=="REGION IV-A CALABARZON", c("Province","CityProvince","Barangay","Population")]
popIVA$den = popIVA$Population / (regionarea[9,2]/length(population$Region[population$Region == "REGION IV-A CALABARZON"]))
popIVA

popARMM = population[population$Region=="AUTONOMOUS REGION IN MUSLIM MINDANAO (ARMM)", c("Province", "CityProvince","Barangay","Population")]
popARMM$den = popARMM$Population / (regionarea[1,2]/length(population$Region[population$Region == "AUTONOMOUS REGION IN MUSLIM MINDANAO (ARMM)"]))
popARMM

popCAR = population[population$Region=="CORDILLERA ADMINISTRATIVE REGION", c("Province", "CityProvince","Barangay","Population")]
popCAR$den = popCAR$Population / (regionarea[2,2]/length(population$Region[population$Region == "CORDILLERA ADMINISTRATIVE REGION"]))
popCAR

popMIMAROPA = population[population$Region=="MIMAROPA REGION", c("Province", "CityProvince","Barangay","Population")]
popMIMAROPA$den = popMIMAROPA$Population / (regionarea[3,2]/length(population$Region[population$Region == "MIMAROPA REGION"]))
popMIMAROPA

popNCR = population[population$Region=="NATIONAL CAPITAL REGION", c("Province", "CityProvince","Barangay","Population")]
popNCR$den = popNCR$Population / (regionarea[4,2]/length(population$Region[population$Region == "NATIONAL CAPITAL REGION"]))
popNCR

popNegros = population[population$Region=="NEGROS ISLAND REGION", c("Province", "CityProvince","Barangay","Population")]
popNegros$den = popNegros$Population / (regionarea[5,2]/length(population$Region[population$Region == "NEGROS ISLAND REGION"]))
popNegros

popI = population[population$Region=="REGION I - ILOCOS", c("Province", "CityProvince","Barangay","Population")]
popI$den = popI$Population / (regionarea[6,2]/length(population$Region[population$Region == "REGION I - ILOCOS"]))
popI

popII = population[population$Region=="REGION II - CAGAYAN VALLEY", c("Province", "CityProvince","Barangay","Population")]
popII$den = popII$Population / (regionarea[7,2]/length(population$Region[population$Region == "REGION II - CAGAYAN VALLEY"]))
popII

popIII = population[population$Region=="REGION III - CENTRAL LUZON", c("Province", "CityProvince","Barangay","Population")]
popIII$den = popIII$Population / (regionarea[8,2]/length(population$Region[population$Region == "REGION III - CENTRAL LUZON"]))
popIII
```

#Merging
The followng lines of codes merge all the regions in a single table using rbind then sorted using order. The head function was used to identify the top five baranagays with the highest population density
```{r include=FALSE}
merge <- rbind(popARMM,popCAR,popMIMAROPA,popNCR,popNegros,popI,popII,popIII,popIVA, popV, popVI, popVII, popVIII, popIX, popX, popXI, popXII, popXIII)
merge
mergesort = merge[order(merge$den, decreasing=TRUE),]
mergetop5 = head(mergesort, 5)
mergetop5
```

#WriteCsv
The following line of code was used to generate a csv file of the top five baranagays with the highest population density
```{r include=FALSE}
write.csv(mergetop5, file="BarangayDensity.csv",row.names=FALSE)
```
